Wait-free Queues with Polygarithmic Step Complexity

### 2022-03-25 article/psuedocode.tex
complete garbage collection methods

\Statex $\blacktriangleright$ \texttt{\textsl{Node}}
\begin{itemize}
\item \texttt{\textsl{*Node} left, right, parent}
\item \texttt{\textsl{Block[]} blocks}\textsf{: index 0 contains an empty block with all fields equal to 0 and \texttt{en} pointers to the first block of the coresponding children. \texttt{blocks[i]} returns the $i$th block stored.}
\item \texttt{\textsl{int} head= 1}\textsf{: index of the first empty cell of \texttt{blocks}}
\item \texttt{\textsl{int} counter= 0}\textsf{}
\item \texttt{\textsl{int[]} super}\textsf{: \texttt{super[i]} stores the index of a superblock in parent that contains some block of this node whose \texttt{time} is field \texttt{i}}
\end{itemize}

@@

\Function{Object}{Dequeue()}{}
\State \texttt{block b= \Call{new}{\textsl{block}}}
\State \texttt{b.element= null}
\State \texttt{b.sum\textsubscript{deq}=1}
\State \Call{Append}{b}
\State \texttt{<i, b>=} \Call{Index}{l\textsubscript{pid}, b.order, 1}
\State \texttt{res=} \Call{ComputeHead}{i, b} \Comment{Index of the enqueue whose argument should be returned} \label{deqRest}
\State \Return \Call{Get}{res}
\State \texttt{b\textsubscript{i}= b}\Comment{block in the root contains the invocation of dequeue}
\State \texttt{b\textsubscript{r}= root.blocks.get(sum\textsubscript{enq}==i)}\Comment{block in the root contains the invocation of dequeue}
\State \texttt{b\textsubscript{i}.age= b\textsubscript{i}.age+1}
\State \texttt{b\textsubscript{r}.age= b\textsubscript{r}.age+1}
\If{\texttt{b\textsubscript{i} or b\textsubscript{r} become old}} \texttt{update maxOld}
\EndIf
\EndFunction{Dequeue}

@@

\If{\texttt{n.parent} \textbf{is} \texttt{null}}
\State \texttt{b.size= max(root.blocks[i-1].size + b.num\textsubscript{enq} - b.num\textsubscript{deq}, 0)}
\State \texttt{b.sum\textsubscript{non-null deq}= root.blocks[i-1].sum\textsubscript{non-null deq} + max( b.num\textsubscript{deq} - root.blocks[i-1].size - b.num\textsubscript{enq}, 0)}
\EndIf

@@

\Statex $\leadsto$ \textsf{Precondition:~\texttt{n.blocks[b]} contains $\geq$\texttt{i} enqueues.}
\Function{element}{Get}{\textsl{node} n, \textsl{int} b, \textsl{int} i} 
\Comment{\textmd{Returns the \texttt{i}th Enqueue in \texttt{b}th block of node \texttt{n}}}
\If{\texttt{n {\keywordfont is} leaf}} \Return \texttt{n.ops[b]}
\Else
\If{\texttt{i $\leq$ n.blocks[b].num\textsubscript{enq-left}}} \Comment{\texttt{i} exists in the left child of \texttt{n}}
\State \texttt{subBlock= \Call{BSearch}{n.left, sum\textsubscript{enq}, i, n.blocks[b-1].end\textsubscript{left}+1, n.blocks[b].end\textsubscript{left}}}
\State \Return\Call{Get}{n.left, subBlock, i-n.left.blocks[subBlock-1].sum\textsubscript{enq}} 
\Else
\State \texttt{i= i-n.blocks[b].num\textsubscript{enq-left}}
\State\texttt{subBlock=\Call{BSearch}{n.right, sum\textsubscript{enq}, i, n.blocks[b-1].end\textsubscript{right}+1, n.blocks[b].end\textsubscript{right}}}
\State \Return\Call{Get}{n.right, subBlock, i-n.right.blocks[subBlock-1].sum\textsubscript{enq}} 
\EndIf
\EndIf
\EndFunction{Get}

@@

\Function{void}{Help}{}\Comment{Helps pending operations}
\For{\texttt{leaf l} \textbf{in leaves}}\Comment{\textit{how to iterate over them?}}
\State{\texttt{last= l.head-1}}
\If{\texttt{l.blocks[last]} \textbf{is not} \texttt{null}}
\If{\texttt{l.blocks[last].element==null}} \Comment{operation is dequeue}
\State \texttt{goto \ref{deqRest} with} \Comment{run \texttt{Dequeue()}  for \texttt{l.ops[last]} after Propagate()}
\State \texttt{l.responses[last]= response}
\EndIf
\EndIf
\EndFor
\EndFunction{Help}
\Statex

@@

\Function{Block}{FindYoungestOld}{b}
\State\Return  \texttt{read all maxOld values among leaves and decide the largest one} \Comment{There is no need to do a sasDK;Lnapgshot.}
\EndFunction{findYoungestOld}

@@

\If{operation i in leaf l cannot find its desired RootBlock}

\State \Return \texttt{l.response[i]}
\EndIf

>>>>

\Statex $\blacktriangleright$ \texttt{\textsl{Node}}
\begin{itemize}
\item \texttt{\textsl{*Node} left, right, parent}
\item \texttt{\textsl{Block[]} blocks}\textsf{: index 0 contains an empty block with all fields equal to 0 and \texttt{en} pointers to the first block of the corresponding children. \texttt{blocks[i]} returns the $i$th block stored.  In the root node it is implemented with a persistent red-black tree and it is a big array in other nodes.}
\item \texttt{\textsl{int} head= 1}\textsf{: index of the first empty cell of \texttt{blocks}}
\item \texttt{\textsl{int} counter= 0}\textsf{}
\item \texttt{\textsl{int[]} super}\textsf{: \texttt{super[i]} stores the index of a superblock in parent that contains some block of this node whose \texttt{time} is field \texttt{i}}
\end{itemize}

@@

\Function{Object}{Dequeue()}{}
\State \texttt{block b= \Call{new}{\textsl{block}}}
\State \texttt{b.element= null}
\State \texttt{b.sum\textsubscript{deq}=1}
\State \Call{Append}{b}
\State \texttt{<i, b>=} \Call{Index}{l\textsubscript{pid}, b.order, 1}
\State \texttt{res=} \Call{ComputeHead}{i, b} \Comment{Index of the enqueue whose argument should be returned} \label{deqRest}
\State \Return \Call{Get}{res}
\State \texttt{b\textsubscript{i}= b}\Comment{block in the root contains the invocation of dequeue}
\State \texttt{b\textsubscript{r}= root.blocks.get(sum\textsubscript{enq}==i)}\Comment{block in the root contains the invocation of dequeue}
\State \texttt{b\textsubscript{i}.age= b\textsubscript{i}.age+1}
\State \texttt{b\textsubscript{r}.age= b\textsubscript{r}.age+1}
\If{\texttt{b\textsubscript{i}==b\textsubscript{i}.num or b\textsubscript{r}==b.num}}\Comment{become old}
\State \texttt{this.leaf.maxOld= Max(b\textsubscript{i}*(b\textsubscript{i}-b\textsubscript{i}.num), b\textsubscript{r}*(b\textsubscript{r}-b\textsubscript{r}.num), this.leaf.maxold)}
\EndIf
\EndFunction{Dequeue}

@@

\If{\texttt{n.parent} \textbf{is} \texttt{null}}
\State \texttt{Cast(b, RootBlock)} \Comment{cast block to a root block}
\State \texttt{b.size= max(root.blocks[i-1].size + b.num\textsubscript{enq} - b.num\textsubscript{deq}, 0)}
\State \texttt{b.sum\textsubscript{non-null deq}= root.blocks[i-1].sum\textsubscript{non-null deq} + max( b.num\textsubscript{deq} - root.blocks[i-1].size - b.num\textsubscript{enq}, 0)}
\EndIf

@@

\Statex $\leadsto$ \textsf{Precondition:~\texttt{n.blocks[b]} contains $\geq$\texttt{i} enqueues.}
\Function{element}{Get}{\textsl{node} n, \textsl{int} b, \textsl{int} i} 
\Comment{\textmd{Returns the \texttt{i}th Enqueue in \texttt{b}th block of node \texttt{n}}}
\If{\texttt{n {\keywordfont is} leaf}} \Return \texttt{n.blocks[b].element}
\Else
\If{\texttt{i $\leq$ n.blocks[b].num\textsubscript{enq-left}}} \Comment{\texttt{i} exists in the left child of \texttt{n}}
\State \texttt{subBlock= \Call{BSearch}{n.left, sum\textsubscript{enq}, i, n.blocks[b-1].end\textsubscript{left}+1, n.blocks[b].end\textsubscript{left}}}
\State \Return\Call{Get}{n.left, subBlock, i-n.left.blocks[subBlock-1].sum\textsubscript{enq}} 
\Else
\State \texttt{i= i-n.blocks[b].num\textsubscript{enq-left}}
\State\texttt{subBlock=\Call{BSearch}{n.right, sum\textsubscript{enq}, i, n.blocks[b-1].end\textsubscript{right}+1, n.blocks[b].end\textsubscript{right}}}
\State \Return\Call{Get}{n.right, subBlock, i-n.right.blocks[subBlock-1].sum\textsubscript{enq}} 
\EndIf
\EndIf
\EndFunction{Get}

@@

\Function{void}{Help}{}\Comment{Helps pending operations}
\For{\texttt{leaf l} \textbf{in leaves}}\Comment{\textit{how to iterate over them?}}
\State{\texttt{last= l.head-1}}
\If{\texttt{l.blocks[last]} \textbf{is not} \texttt{null}}
\If{\texttt{l.blocks[last].element==null}} \Comment{operation is dequeue}
\State \texttt{goto \ref{deqRest} with these values <>} \Comment{run \texttt{Dequeue()}  for \texttt{l.ops[last]} after Propagate(). \textit{TODO}}
\State \texttt{l.responses[last]= response}
\EndIf
\EndIf
\EndFor
\EndFunction{Help}
\Statex

@@

\Function{Block}{FindYoungestOld}{b}
\For{\texttt{leaf l} \textbf{in leaves}}
\State\texttt{max= Max(l.maxOld, max)}
\EndFor
\State\Return\texttt{max} \Comment{This snapshot suffies.}
\EndFunction{findYoungestOld}

@@

\If{a dequeue cannot find the root block}

\State \Return \texttt{this.leaf.response(block.order)}
\EndIf
